---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(8675309)
```

# Data Simulation Workshop Materials

You can install the packages used in these tutorials and get a function that makes it easy to access the workshop .Rmd files by running the following code:

```{r, eval = FALSE}
devtools::install_github("debruine/data-sim-workshops")
```

Load exercises with the following code:

```{r, eval = FALSE}
dsw::exercise("faux")
dsw::exercise("calories")
dsw::exercise("fixed")
dsw::exercise("mixed")
```



## Upcoming Workshops

### Data simulation for factorial designs

When: Tuesday July 13 13:30-15:00 (BST)  
Where: From Proposal to Publication: Pathways to Open Science (zoom)

This session will cover the basics of simulation using {faux}. We will simulate data with factorial designs by specifying the within and between-subjects factor structure, each cell mean and standard deviation, and correlations between cells where appropriate. This can be used to create simulated data sets to be used in preparing the analysis code for pre-registrations or registered reports. We will also create data sets for simulation-based power analyses. Students will need to have very basic knowledge of R and R Markdown, and have installed {faux}, {afex} and {tidyverse}.

#### Prep

* Install R packages from CRAN: `tidyverse`, `afex`, `faux`, and `broom`
* Download files:  [faux-stub.Rmd](https://raw.githubusercontent.com/debruine/data-sim-workshops/master/inst/stubs/faux-stub.Rmd) &
[calories-stub.Rmd](https://raw.githubusercontent.com/debruine/data-sim-workshops/master/inst/stubs/calories-stub.Rmd)

### Data simulation for mixed designs

When: Tuesday July 13 15:30-17:00 (BST)  
Where: From Proposal to Publication: Pathways to Open Science (zoom)

This session will cover simulating data for a mixed design, where trials are crossed with subjects. We will learn how to analyse this using {lme4}, with a focus on understanding how the simulation parameters correspond to the output. Finally, we will learn how to use simulation to calculate power. Students will need to have basic knowledge of R and R Markdown, some familiarity with mixed designs (even if they don't currently analyse them with mixed models) and have installed {faux}, {afex}, {tidyverse}, and {lme4}.

#### Prep

* Install R packages from CRAN: `tidyverse`, `afex`, `lme4`, `broom`, `broom.mixed`, `faux`
* Download files: [mixed-stub.Rmd](https://raw.githubusercontent.com/debruine/data-sim-workshops/master/inst/stubs/mixed-stub.Rmd)


## Resources

* [Data Skills for Reproducible Science](https://psyteachr.github.io/msc-data-skills/) open source textbook introducing tidyverse for psychologists
* [Understanding mixed effects models through data simulation](https://osf.io/3cz2e/) (preprint, code, and shiny apps) 
* [Simulate Basic Distributions](https://shiny.psy.gla.ac.uk/debruine/simulate/)
* [Scienceverse](http://shiny.ieis.tue.nl/scienceverse/) (shiny app has an interactive data simulation tab)

## Past Workshops

* University of Glasgow
    Institute of Neuroscience and Psychology 
    2020 Jan 28 13:00-15:00 and Feb 5 14:00-16:00 
    
* University of Grenoble
    Understanding Mixed-Effects Models through Data Simulation
    2021 February 5 13:00-15:00 

* PsyPAG Data Simulation Summer School 
    Simulation for factorial designs with faux
    2021 June 4 13:00-15:00 


