---
title: "Intro to Faux"
author: "Lisa DeBruine"
output: 
  rmarkdown::html_vignette:
    df_print: kable
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  fig.width = 5,
  fig.height = 3,
  dpi = 144
)
set.seed(8675309) # Jenny, I've got your number
```

```{r libs, message=FALSE}
library(tidyverse)
library(faux)
library(broom)
library(afex)
```

In this tutorial, we'll learn how to simulate data for factorial designs using {faux}. There are more extensive examples at <https://debruine.github.io/faux/>.

## Multivariate normal

You can create sets of correlated normally distributed values using `rnorm_multi()`.

```{r}
dat3 <- rnorm_multi(
  n = 50,
  vars = 3,
  mu = c(1, 2, 3),
  sd = c(0.5, 1, 1.5),
  r = c(0, .25, .5),
  varnames = c("A", "B", "C")
)
```

The function `get_params()` gives you a quick way to see the means, SDs and correlations in the simulated data set to make sure you set the parameters correctly.

```{r}
get_params(dat3)
```

If you set `empirical` to `TRUE`, the values you set will be the **sample** parameters, not the **population** parameters. This isn't usually what you want for a simulation, but can be useful to check you set the parameters correctly.

```{r}
dat3 <- rnorm_multi(
  n = 50,
  vars = 3,
  mu = c(1, 2, 3),
  sd = c(0.5, 1, 1.5),
  r = c(0, .25, .5),
  varnames = c("A", "B", "C"),
  empirical = TRUE
)

get_params(dat3)
```

### Shortcuts

There are a few shortcuts you can use. Run the following and see if you can guess how they work.

```{r}
guess1 <- rnorm_multi(50, mu = c(x = 1, y = 2, z = 3), empirical = TRUE)

get_params(guess1)
```

```{r}
guess2 <- rnorm_multi(50, vars = 4, r = 0.5, empirical = TRUE)

get_params(guess2)
```

```{r}
iris_r <- cor(iris[, 1:4])
iris_mu <- summarise_all(iris[, 1:4], mean) %>% t()
iris_sd <- summarise_all(iris[, 1:4], sd) %>% t()

guess3 <- rnorm_multi(50, 
                      mu = iris_mu, 
                      sd = iris_sd, 
                      r = iris_r)

get_params(guess3)
```

You can set the r for correlations in a few different ways.

If all correlations have the same value, just set r equal to a single number.

```{r}
# all correlations the same value
rho_same <- rnorm_multi(50, 4, r = .5, empirical = TRUE)
get_params(rho_same)
```

You can set rho to a vector or matrix of the full correlation matrix. This is convenient when you're getting the values from an existing dataset, where you can just use the output of the `cor()` function.

```{r}
# full correlation matrix
#         X1   X2   X3   X4
rho <- c(1.0, 0.5, 0.4, 0.3, # X1
         0.5, 1.0, 0.2, 0.1, # X2
         0.4, 0.2, 1.0, 0.0, # X3
         0.3, 0.1, 0.0, 1.0) # X4
rho_cormat <- rnorm_multi(50, 4, r = rho, empirical = TRUE)
get_params(rho_cormat)
```

Alternatively, you can just specify the values from the upper right triangle of a correlation matrix. This is easier if you're reading the values out of a paper.

```{r}
# upper right triangle
#         X2   X3   X4
rho <- c(0.5, 0.4, 0.3, # X1
              0.2, 0.1, # X2
                   0.0) # X3

rho_urt <- rnorm_multi(50, 4, r = rho, empirical = TRUE)
get_params(rho_urt)
```


## Factorial Designs

You can use `rnorm_multi()` to simulate data for each between-subjects cell of a factorial design and manually combine the tables, but faux has a function that better maps onto how we usually think and teach about factorial designs.

The default design is 100 observations of one variable (named `y`) with a mean of 0 and SD of 1. Unless you set `plot = FALSE` or run `faux_options(plot = FALSE)`, this function will show you a plot of your design so you can check that it looks like you expect.

```{r}
simdat1 <- sim_design()
```


### Factors

Use lists to set the names and levels of `within`- and `between`-subject factors.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret"))
)
```

You can set `mu` and `sd` with unnamed vectors, but getting the order right can take some trial and error.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  mu = 1:6
)
```

You can set values with a named vector for a single type of factor. The values do not have to be in the right order if they're named.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  mu = c(cat = 1, ferret = 5, dog = 3),
  sd = c(pre = 1, post = 2)
)
```

Or use a data frame for within- and between-subject factors.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  mu = data.frame(
    pre = c(1, 3, 5),
    post = c(2, 4, 6),
    row.names = c("cat", "dog", "ferret")
  )
)
```

If you have within-subject factors, set the correlations for each between-subject cell like this.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  r = list(cat = 0.5,
           dog = 0.25,
           ferret = 0),
  empirical = TRUE,
  plot = FALSE
)

get_params(pettime)
```

You can also change the name of the `dv` and `id` columns and output the data in long format. If you do this, you also need to tell `get_params()` what columns contain the between- and within-subject factors, the dv, and the id.

```{r}
dat_long <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  id = "subj_id",
  dv = "score",
  long = TRUE,
  plot = FALSE
)

get_params(dat_long, 
           between = "pet", 
           within = "time",
           id = "subj_id",
           dv = "score",
           digits = 3)
```

*The current version of faux doesn't actually need `between` or `dv` and later versions won't need `within` or `id` if a data frame was created with faux.*

### Anonymous Factors

If you need to make a quick demo, you can set factors anonymously with integer vectors. For example, the following code makes 3B\*2B\*2W mixed design.

```{r}
dat_anon <- sim_design(
  n = 50,
  between = c(3, 2),
  within = 2,
  mu = 1:12
)
```

Faux has a quick plotting function for visualising data made with faux. The plot created by `sim_design()` shows the *design*, while this function shows the simulated *data*.

```{r}
plot(dat_anon)
```
You can change the order of plotting and the types of geoms plotted. This takes a little trial and error, so this function will probably be refined in later versions.

```{r}
plot(dat_anon, "B", "A", "C", geoms = c("violin", "pointrangeSD"))
```

## Replications

You often want to simulate data repeatedly to do things like calculate power. The `sim_design()` function has a lot of overhead for checking if a design makes sense and if the correlation matrix is possible, so you can speed up the creation of multiple datasets with the same design using the `rep` argument. This will give you a nested data frame with each dataset in the `data` column. 

```{r}
dat_rep <- sim_design(
  within = 2,
  n = 20,
  mu = c(0, 0.25),
  rep = 5,
  plot = FALSE
)
```

### Analyse each replicate

You can run analyses on the nested data by wrapping your analysis code in a function then using `map()` to run the analysis on each data set and `unnest()` to expand the results into a data table.

```{r}
analyse <- function(data) {
  t.test(data$A1, data$A2, paired = TRUE) %>% broom::tidy()
}

dat_rep %>%
  mutate(analysis = map(data, analyse)) %>%
  select(-data) %>%
  unnest(analysis)
```

### ANOVA

Use the same pattern to run an ANOVA on a version of the `pettime` dataset.

First, simulate 100 datasets in long format. These data will have small main effects of pet and time, but no interaction.

```{r}
pettime100 <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog")),
  n = c(cat = 50, dog = 40),
  mu = data.frame(
    pre = c(1, 1.2),
    post = c(1.2, 1.4),
    row.names = c("cat", "dog")
  ),
  sd = 1,
  id = "pet_id",
  dv = "score",
  r = 0.5,
  long = TRUE,
  rep = 100
)
```

Then set up your analysis. We'll use the `aov_ez()` function from the {afex} package because its arguments match those of `sim_design()`.

```{r}
afex::set_sum_contrasts() # avoids annoying afex message
afex_options(include_aov = FALSE) # runs faster
afex_options(es_aov = "pes") # changes effect size measure to partial eta squared

analyse <- function(data) {
  a <- afex::aov_ez(
    id = "pet_id",
    dv = "score",
    between = "pet",
    within = "time",
    data = data
  ) 
  # return anova_table for GG-corrected DF
  as_tibble(a$anova_table, rownames = "term") %>%
    mutate(term = factor(term, levels = term)) %>% # keeps terms in order
    rename(p.value = `Pr(>F)`) # fixes annoying p.value name
}
```

Make a table of the results of each analysis:

```{r}
pettime_sim <- pettime100 %>%
  mutate(analysis = map(data, analyse)) %>%
  select(-data) %>%
  unnest(analysis)
```

```{r, echo = FALSE}
# show the first 6 rows
head(pettime_sim) %>% 
  mutate(across(5:8, round, 3))
```

Then you can summarise the data to calculate things like power for each effect or mean effect size.

```{r}
pettime_sim %>%
  group_by(term) %>%
  summarise(power = mean(p.value < 0.05),
            mean_pes = mean(pes),
            .groups = "drop")
```

The power for the between-subjects effect of pet is smaller than for the within-subjects effect of time. What happens if you reduce the correlation between pre and post?

## Exercises

### Multivariate normal

Sample 40 values of three variables named `J`, `K` and `L` from a population with means of 10, 20 and 30, and SDs of 5. `J` and `K` are correlated 0.5, `J` and `L` are correlated 0.25, and `K` and `L` are not correlated.

```{r, include=FALSE}
ex1 <- rnorm_multi(n = 40, mu = c(J = 10, K = 20, L = 30),
                   sd = 5, r = c(0.5, 0.25, 0))

get_params(ex1)
```

### From existing data

Using the data from the built-in dataset `attitude`, simulate a new set of 20 observations drawn from a population with the same means, SDs and correlations for each column as the original data.

```{r, include=FALSE}
dat_r <- cor(attitude)
dat_mu <- summarise_all(attitude, mean) %>% t()
dat_sd <- summarise_all(attitude, sd) %>% t()

ex2 <- rnorm_multi(20, mu = dat_mu, sd = dat_sd,r = dat_r)

get_params(ex2)
```


### 2b

Create a dataset with a between-subject factor of "pet" having two levels, "cat", and "dog". The DV is "happiness" score. There are 20 cat-owners with a mean happiness score of 10 (SD = 3) and there are 30 dog-owners with a mean happiness score of 11 (SD = 3).

```{r, include=FALSE}
dat2b <- sim_design(
  between = list(pet = c("cat", "dog")),
  dv = "happiness",
  n = list(cat = 20, dog = 30),
  mu = list(cat = 10, dog = 11),
  sd = 3
)

get_params(dat2b, between = "pet")
```

### 3w

Create a dataset of 20 observations with 1 within-subject variable ("condition") having 3 levels ("A", "B", "C") with means of 10, 20 and 30 and SD of 5. The correlations between each level have r = 0.4. The dataset should look like this:

| id | condition | score |
|:---|:----------|------:|
|S01 | A         |  9.17 |
|... | ...       |  ...  |
|S20 | A         | 11.57 |
|S01 | B         | 18.44 |
|... | ...       |  ...  |
|S20 | B         | 20.04 |
|S01 | C         | 35.11 |
|... | ...       |  ...  |
|S20 | C         | 29.16 |

```{r, include=FALSE}

dat3w <- sim_design(
  within = list(condition = c("A", "B", "C")),
  n = 20,
  mu = c(10, 20, 30),
  sd = 5,
  r = .4,
  dv = "score",
  long = TRUE
)

get_params(dat3w)

```

### 2w*2w

Create a dataset with 50 observations of 2 within-subject variables ("A" and "B") each having 2 levels. The mean for all cells is 10 and the SD is 2. The dataset should have 20 subjects. The correlations look like this:

|       | A1_B1 | A1_B2 | A2_B1 | A2_B2 |
|:------|------:|------:|------:|------:|
| A1_B1 | 1.0   | 0.5   | 0.5   | 0.2   |
| A1_B2 | 0.5   | 1.0   | 0.2   | 0.5   |
| A2_B1 | 0.5   | 0.2   | 1.0   | 0.5   |
| A2_B2 | 0.2   | 0.5   | 0.5   | 1.0   |


```{r, include=FALSE}
dat2w2w <- sim_design(
  within = c(2,2),
  n = 50, 
  mu = 10,
  sd = 2,
  r = c(.5, .5, .2, 
            .2, .5, 
                .5)
)

get_params(dat2w2w)
```

### 2w*3b

Create a dataset with a between-subject factor of "pet" having 3 levels ("cat", "dog", and "ferret") and a within-subject factor of "time" having 2 levels ("pre" and "post"). The N in each group should be 10. Means are:

* cats: pre = 10, post = 12
* dogs: pre = 14, post = 16
* ferrets: pre = 18, post = 20

SDs are all 5 and within-cell correlations are all 0.25.

```{r, include=FALSE}

mu <- data.frame(
  cat = c(10, 12),
  dog = c(14, 16),
  ferret = c(18, 20)
)

dat2w3b <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  n = 10,
  mu = mu,
  sd = 5,
  r = 0.25
)

get_params(dat2w3b)
  
```

### Replications

Create 5 datasets with a 2b*2b design, 30 participants in each cell. Each cell's mean should be 0, except A1_B1, which should be 0.5. The SD should be 1. Make the resulting data in long format.

```{r, include=FALSE}
dat2b2b <- sim_design(
  between = c(2,2),
  n = 30,
  mu = c(0.5, 0, 0, 0),
  rep = 5,
  long = TRUE
)
```

### Power 

Simulate 100 datasets like the one above and use `lm()` or `afex::aov_ez()` to look at the interaction between A and B. What is the power of this design?

```{r, include=TRUE}
# simulate 100 datasets
dat2b2b_100 <- sim_design(
  between = c(2, 2),
  n = 30,
  mu = c(0.5, 0, 0, 0),
  rep = 100,
  long = TRUE
)

# linear model version
analyse_lm <- function(data) {
  lm(y ~ A*B, data = data) %>% 
    broom::tidy() %>%
    mutate(term = factor(term, levels = term))
}

ana_lm <- dat2b2b_100 %>%
  mutate(analysis = map(data, analyse_lm)) %>%
  select(-data) %>%
  unnest(analysis)

ana_lm %>%
  group_by(term) %>%
  summarise(power = mean(p.value < .05),
            .groups = "drop")

# anova version
analyse_aov <- function(data) {
  a <- afex::aov_ez(id = "id",
              dv = "y",
              between = c("A", "B"),
              data = data)
  
  as_tibble(a$anova_table, rownames = "term") %>%
    mutate(term = factor(term, levels = term)) %>% 
    rename(p.value = `Pr(>F)`) 
}

ana_aov <- dat2b2b_100 %>%
  mutate(analysis = map(data, analyse_aov)) %>%
  select(-data) %>%
  unnest(analysis)

ana_aov %>%
  group_by(term) %>%
  summarise(power = mean(p.value < .05),
            .groups = "drop")
```


