---
title: "Intro to Faux"
output: 
  html_document:
    df_print: kable
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%",
  fig.width = 5,
  fig.height = 3,
  dpi = 144
)
```

```{r libs, message=FALSE}
library(tidyverse)
library(faux)
library(broom)
library(afex)
```

In this tutorial, we'll learn how to simulate data for factorial designs using {faux}. There are more extensive examples at <https://debruine.github.io/faux/>.

## Multivariate normal

You can create sets of correlated normally distributed values using `rnorm_multi()`.

```{r}
dat3 <- rnorm_multi(
  n = 50,
  vars = 3,
  mu = c(1, 2, 3),
  sd = c(0.5, 1, 1.5),
  r = c(0, .25, .5),
  varnames = c("A", "B", "C")
)
```

The function `get_params()` gives you a quick way to see the means, SDs and correlations in the simulated data set to make sure you set the parameters correctly.

```{r}
get_params(dat3)
```

If you set `empirical` to `TRUE`, the values you set will be the **sample** parameters, not the **population** parameters. This isn't usually what you want for a simulation, but can be useful to check you set the simulation parameters correctly.

```{r}
dat3 <- rnorm_multi(
  n = 50,
  vars = 3,
  mu = c(1, 2, 3),
  sd = c(0.5, 1, 1.5),
  r = c(0, .25, .5),
  varnames = c("A", "B", "C"),
  empirical = TRUE
)

get_params(dat3)
```

### Shortcuts

There are a few shortcuts you can use. Run the following and see if you can guess how they work.

```{r}
guess1 <- rnorm_multi(50, mu = c(x = 1, y = 2, z = 3), empirical = TRUE)

get_params(guess1)
```

```{r}
guess2 <- rnorm_multi(50, vars = 4, r = 0.5, empirical = TRUE)

get_params(guess2)
```

```{r}
iris_r <- cor(iris[, 1:4])
iris_mu <- summarise_all(iris[, 1:4], mean) %>% t()
iris_sd <- summarise_all(iris[, 1:4], sd) %>% t()

guess3 <- rnorm_multi(50, 
                      mu = iris_mu, 
                      sd = iris_sd, 
                      r = iris_r)

get_params(guess3)
```

You can set the r for correlations is a few different ways.

```{r}
# all correlations the same value
rho_same <- rnorm_multi(50, 4, r = .5, empirical = TRUE)
get_params(rho_same)
```

```{r}
# upper right triangle
rho_urt <- rnorm_multi(50, 4, 
                      #      X2   X3   X4
                       r = c(0.5, 0.4, 0.3, # X1
                                  0.2, 0.1, # X2
                                      0.0), # X3
                       empirical = TRUE)
get_params(rho_urt)
```

```{r}
# full correlation matrix
rho_cormat <- rnorm_multi(50, 4, 
                        #      X1   X2   X3   X4
                          r = c(1.0, 0.5, 0.4, 0.3,  # X1
                                0.5, 1.0, 0.2, 0.1,  # X2
                                0.4, 0.2, 1.0, 0.0,  # X3
                                0.3, 0.1, 0.0, 1.0), # X4
                       empirical = TRUE)
get_params(rho_cormat)
```


## Factorial Designs

You can just use `rnorm_multi()` to simulate data for each between-subjects cell of a factorial design and manually combine the tables, but faux has a function that better maps onto how we usually think and teach about factorial designs.

The default design is 100 observations of one variable (named `y`) with a mean of 0 and SD of 1. Unless you set `plot = FALSE` or run `faux_options(plot = FALSE)`, this function will show you a plot of your design so you can check that it looks like you expect.

```{r}
simdat1 <- sim_design()
```


### Factors

Use lists to set the names and levels of within- and between-subject factors.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret"))
)
```

You can set mu and sd with unnamed vectors, but getting the order right can be tricky.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  mu = 1:6
)
```

You can set values with a named vector for a single type of factor. The values do not have to be in the right order if they're named.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  mu = c(cat = 1, ferret = 5, dog = 3),
  sd = c(pre = 1, post = 2)
)
```

Or use a data frame for within- and between-subject factors.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  mu = data.frame(
    pre = c(1, 3, 5),
    post = c(2, 4, 6),
    row.names = c("cat", "dog", "ferret")
  )
)
```

If you have within-subject factors, set the correlations for each between-subject cell like this. You need to tell `get_params()` if you have any between-subject columns.

```{r}
pettime <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  r = list(cat = 0.5,
           dog = 0.25,
           ferret = 0),
  empirical = TRUE,
  plot = FALSE
)

get_params(pettime, between = "pet")
```

You can also change the name of the `dv` and `id` columns and output the data in long format. If you do this, you also need to tell `get_params()` what columns contain the between- and within-subject factors, the dv, and the id. 

```{r}
dat_long <- sim_design(
  within = list(time = c("pre", "post")),
  between = list(pet = c("cat", "dog", "ferret")),
  id = "subj_id",
  dv = "score",
  long = TRUE,
  plot = FALSE
)

get_params(dat_long, 
           between = "pet", 
           within = "time",
           id = "subj_id",
           dv = "score",
           digits = 3)
```

### Anonymous Factors

If you need to make a quick demo, you can set factors anonymously with integer vectors.

```{r}
dat_anon <- sim_design(
  n = 50,
  between = c(3, 2),
  within = 2,
  mu = 1:12
)
```

Faux has a quick plotting function for visualising data made with sim_design.

```{r}
plot(dat_anon)
```
You can change the order of plotting and the types of geoms plotted. This takes a little trial and error, so this function will probably be refined in later versions.

```{r}
plot(dat_anon, "B", "A", "C", geoms = c("violin", "pointrangeSD"))
```

### Replications

You often want to simulate data repeatedly to do things like calculate power. The `sim_design()` function has a lot of overhead for checking if a design makes sense and if the correlation matrix is possible, so you can speed up the creation of multiple datasets with the same design using the `rep` argument. This will give you a nested data from with each dataset in the `data` column. 

```{r}
dat_rep <- sim_design(
  within = 2,
  n = 20,
  mu = c(0, 0.25),
  rep = 10,
  plot = FALSE
)
```

You can run analyses on the nested data like this:

```{r}
map_df(dat_rep$data, ~{
 t.test(.x$A1, .x$A2, paired = TRUE) %>% broom::tidy()
})
```


## Exercises

### Multivariate normal

Sample 40 values of three variables named `J`, `K` and `L` from a population with means of 10, 20 and 30, and SDs of 5. `J` and `K` are correlated 0.5, `J` and `L` are correlated 0.25, and `K` and `L` are not correlated.

```{r}

```

### From existing data

Using the data from the built-in dataset `attitude`, simulate a new set of 20 observations drawn from a population with the same means, SDs and correlations for each column as the original data.

```{r}

```


### 2b

Create a dataset with a between-subject factor of "pet" having two levels, "cat", and "dog". The DV is "happiness" score. There are 20 cat-owners with a mean happiness score of 10 (SD = 3) and there are 30 dog-owners with a mean happiness score of 11 (SD = 3).

```{r}

```

### 3w

Create a dataset of 20 observations with 1 within-subject variable ("condition") having 3 levels ("A", "B", "C") with means of 10, 20 and 30 and SD of 5. The correlations between each level have r = 0.4. The dataset should look like this:

| id | condition | score |
|:---|:----------|------:|
|S01 | A         |  9.17 |
|... | ...       |  ...  |
|S20 | A         | 11.57 |
|S01 | B         | 18.44 |
|... | ...       |  ...  |
|S20 | B         | 20.04 |
|S01 | C         | 35.11 |
|... | ...       |  ...  |
|S20 | C         | 29.16 |

```{r}

```

### 2w*2w

Create a dataset with 50 observations of 2 within-subject variables ("A" and "B") each having 2 levels. The mean for all cells is 10 and the SD is 2. The dataset should have 20 subjects. The correlations look like this:

|       | A1_B1 | A1_B2 | A2_B1 | A2_B2 |
|:------|------:|------:|------:|------:|
| A1_B1 | 1.0   | 0.5   | 0.5   | 0.2   |
| A1_B2 | 0.5   | 1.0   | 0.2   | 0.5   |
| A2_B1 | 0.5   | 0.2   | 1.0   | 0.5   |
| A2_B2 | 0.2   | 0.5   | 0.5   | 1.0   |


```{r}

```

### 2w*3b

Create a dataset with a between-subject factor of "pet" having 3 levels ("cat", "dog", and "ferret") and a within-subject factor of "time" having 2 levels ("pre" and "post"). The N in each group should be 10. Means are:

* cats: pre = 10, post = 12
* dogs: pre = 14, post = 16
* ferrets: pre = 18, post = 20

SDs are all 5 and within-cell correlations are all 0.25.

```{r}

```

### Replications

Create 5 datasets with a 2b*2b design, 30 participants in each cell. Each cell's mean should be 0, except A1_B1, which should be 0.5. The SD should be 1. Make the resulting data in long format.

```{r}

```

### Power 

Simulate 100 datasets like the one above and use `lm()` or `afex::aov_ez()` to look at the interaction between A and B. What is the power of this design?

```{r}

```

